---
title: "Example"
output: html_document
---

### Get Started

To get started, please load the #Simulator Name# package

```{r, message=FALSE}
library(Seurat)
library(SingleCellExperiment)
library(scuttle)
library(scater)
library(ggplot2)
library(energy)
library(funfun)
library(data.table)
library(cowplot)
library(corrplot)
library(ggsci)
library(egg)
library(expm)
library(BASS)
library(STdeconvolve)
library(SpaceX)
library(BayesSpace)

source("~/Desktop/simulator_YZ.R")
```


### Parameter Setting


```{r}
set.seed(1)
svg_num=50
noise_num=100

x_len <- 100
y_len <- 100
nGenes <- 150


gene_pattern = data.frame(ct1 = c(rep('hotspot',10),rep('streak',10),rep('no_pattern',130)),
                          ct2= c(rep('no_pattern',10),rep('streak',10),rep('hotspot',10),rep('no_pattern',120)),
                          ct3 = c(rep('no_pattern',30),rep('hotspot',10),rep('layer',10),rep('no_pattern',100)))


streak_x <- data.frame(ct1 = c(rep(10,3),rep(20,3),rep(30,4)),
                       ct2 = c(rep(60,3),rep(70,3),rep(80,4)))

hot_spot_center <- list(ct1 = data.frame(x_coord = c(rep(25,10)), y_coord = c(rep(25,3),rep(50,4),rep(75,3))),
                        ct2 = data.frame(x_coord = c(rep(25,4),rep(75,6)), y_coord = c(rep(75,7),rep(25,3))),
                        ct3 = data.frame(x_coord = c(rep(25,3),rep(50,4),rep(75,3)), y_coord = c(rep(75,10))))

gene_relationship = data.table(ct1 = c(rep(NA,10),rep(1,3),rep(2,3),rep(3,4),rep(NA,130)),
                               ct2 = c(rep(NA,10),rep(1,3),rep(2,3),rep(3,4),rep(NA,130)),
                               ct3 = c(rep(NA,40),rep(1,10),rep(NA,100)))


layer_param =  data.frame(ct1 = c(1,1/3,0,0),
                          ct2 = c(0,1/3,1/2,4/5),
                          ct3 = c(0,1/3,1/2,1/5))

# layer_param = data.frame(ct1 = c(1/3,1/3,1/3,1/3),
#                          ct2 = c(1/3,1/3,1/3,1/3),
#                          ct3 = c(1/3,1/3,1/3,1/3))

#define layer
xmin <- c(0, 0, 50, 50)
xmax <- c(50, 50, 100, 100)
ymin <- c(0, 50, 0, 50)
ymax <- c(50, 100, 50, 100)




```

```{r,include=FALSE,message=FALSE}
rownames(gene_pattern) = paste0('gene',1:nGenes)
rownames(gene_relationship) = paste0('gene',1:nGenes)
```


### Simulation

```{r,message=FALSE,warning=FALSE}

coord = generate_coordinates(x_len,y_len,'hex',spot_distance=5)
mean_param = generate_gene_pattern(coord=coord,gene_pattern=gene_pattern,seed=123,
                                   center_coord = hot_spot_center,
                                   streak_x = streak_x)
cov_gene <- generate_covariance_group(gene_relationship)
true_count <- generate_true_count(coord,mean_param,gene_relationship,cov_gene)
layer_coord <- generate_layer(coord, xmin, xmax, ymin, ymax)
celltype_proportion <- generate_cell_prop(layer_coord,layer_param)
#seq_count <- generate_batch_count(true_count, p0 = 0.05, celltype_proportion, seq_depth = 1e6) #optional
spot_count_layers <- generate_spot_count(true_count, nGenes, cell_number = rep(10,nrow(coord)), celltype_proportion)

```

```{r,include=FALSE}
param_all = c()
for(ct in colnames(gene_pattern)){
  temp = mean_param[[ct]] %>% as.data.table()
  temp$cell_type = ct
  param_all = rbind(param_all,temp)
}
coord_param = cbind(coord,param_all)
coord_param = melt(coord_param,id.vars = c('x_coord','y_coord','cell_type'))
colnames(coord_param) <- c("x_coord","y_coord","cell_type","variable","mean")

coord_param |> filter(variable %in% paste0("gene",41:50)) |> ggplot()+geom_point(aes(x=x_coord,y=y_coord,fill=mean),shape=21,size=1)+
  facet_wrap(cell_type~variable,ncol=10)+
  coord_fixed()+xlim(-1,x_len+1)+ylim(-1,y_len+1)+theme_article()+scale_fill_distiller(palette = "Spectral")

prop_all = c()
for(ct in colnames(celltype_proportion)){
  temp = celltype_proportion[,ct] %>% as.data.table()
  prop_all = cbind(prop_all,temp)
  colnames(prop_all)[ncol(prop_all)] <- ct
}
coord_prop = cbind(coord,prop_all)
coord_prop = melt(coord_prop,id.vars = c('x_coord','y_coord'))
colnames(coord_prop) = c('x_coord','y_coord','cell_type','proportion')
ggplot(coord_prop)+geom_point(aes(x=x_coord,y=y_coord,fill=proportion),shape=21,size=2)+
  facet_wrap(as.factor(coord_prop$cell_type),ncol=3)+
  coord_fixed()+xlim(-1,x_len+1)+ylim(-1,y_len+1)+theme_article()+scale_fill_distiller(palette = "Spectral")
layer_factor = as.numeric(as.factor(layer_coord$layer))
```

*Celltype 1 count corr structure*

```{r,echo=FALSE,message=FALSE}
seq_count <- true_count
corrplot(cor(seq_count[["ct1"]][,1:50],method = "spearman"))
```

*Celltype 2 count corr structure*

```{r,echo=FALSE,message=FALSE}
corrplot(cor(seq_count[["ct2"]][,1:50],method = "spearman"))
```

*Celltype 3 count corr structure*

```{r,echo=FALSE,message=FALSE}
corrplot(cor(seq_count[["ct3"]][,1:50],method = "spearman"))
```

*Spot count corr structure*

```{r,echo=FALSE,message=FALSE}
corrplot(cor(spot_count_layers[,1:50],method = "spearman"))
```

*cell-type composition and layer segmentation*

```{r,echo=FALSE,message=FALSE}
# coord_prop = cbind(coord,prop_all)
# coord_prop = melt(coord_prop,id.vars = c('x_coord','y_coord'))
coord_prop_layer <- cbind(coord_prop,layer_factor)
real_prop <- cbind(paste(coord_prop$x_coord,coord_prop$y_coord),coord_prop[,3:4])
prop_plot <- dcast(real_prop, V1 ~ cell_type)
prop_plot <- prop_plot |> mutate(x = as.numeric(sapply(V1,function(i) strsplit(i, "\\s+")[[1]][1])),
                                 y = as.numeric(sapply(V1,function(i) strsplit(i, "\\s+")[[1]][2])))
prop_plot <- prop_plot |> merge(cbind(paste(coord_prop_layer$x_coord,coord_prop_layer$y_coord),coord_prop_layer[,4:5]), by = "V1") |> 
  select(V1, ct1,ct2,ct3,x,y,layer_factor) |> unique()

vizAllTopics(prop_plot[,c("ct1","ct2","ct3")], prop_plot[,c("x","y")],
             groups = prop_plot$layer_factor, 
             group_cols = rainbow(length(unique(layer_factor))),
             r=2,lwd = 0.5)	
```


# Downstream analysis example

### UMAP

```{r,warning=FALSE,message=FALSE}
sce <- SingleCellExperiment(assays = list(counts=t(spot_count_layers)))
colData(sce)$layer = as.factor(layer_coord$layer)
sce <- logNormCounts(sce)
sce <- runPCA(sce, exprs_values = "logcounts",scale = TRUE)
sce <- runUMAP(sce, dimred = "PCA", name = "UMAP")
plotReducedDim(sce, dimred = "UMAP",colour_by = "layer") 
```


### Domain segemntation

#### BASS

```{r}
cnts <- list(t(spot_count_layers))
colnames(cnts[[1]]) <- paste0("spot_",1:ncol(cnts[[1]]))
xy <- list(as.matrix(coord))
rownames(xy[[1]]) <- paste0("spot_",1:ncol(cnts[[1]]))
BASS <- createBASSObject(cnts, xy, C = 3, R = 4, beta_method = "SW",burnin = 10000)

listAllHyper(BASS)
BASS <- BASS.preprocess(BASS,doLogNormalize = TRUE, doPCA = TRUE, scaleFeature = TRUE, nPC = 20,doBatchCorrect = FALSE)
BASS <- BASS.run(BASS)
BASS <- BASS.postprocess(BASS)


BASS_layer <- as.factor(unlist(BASS@results$z))
coord_prop_layer <- cbind(coord_prop,layer_factor,BASS_layer)
ggplot(coord_prop_layer)+geom_point(aes(x=x_coord,y=y_coord,color=BASS_layer),shape=21,size=3) +
  coord_fixed()+xlim(-1,x_len+1)+ylim(-1,y_len+1)+theme_article()+scale_fill_distiller(palette = "Spectral")
```

#### BayesSpace

```{r}
sce_BayesSpace <- sce
row = coord$x
col = coord$y
colData(sce_BayesSpace) = cbind(colData(sce_BayesSpace), row, col)

q <- 4  # Number of clusters
d <- 15  # Number of PCs

sce_BayesSpace <- logNormCounts(sce_BayesSpace)

sce_BayesSpace <- scater::runPCA(sce_BayesSpace)

sce_BayesSpace = spatialPreprocess(sce_BayesSpace, skip.PCA = T, platform = "Visium", log.normalize = F) 
#add BayesSpace metadata, without messing with PCA/logcounts

sce_BayesSpace <- spatialCluster(sce_BayesSpace, q=q, d=d, platform = "Visium",
                        nrep=50000, gamma=3, save.chain=TRUE, model = "t")


BayesSpace.labels <- as.factor(sce_BayesSpace$spatial.cluster)

coord_prop_layer <- cbind(coord_prop,layer_factor,BayesSpace.labels)
ggplot(coord_prop_layer)+geom_point(aes(x=x_coord,y=y_coord,color=BayesSpace.labels),shape=21,size=3) +
  coord_fixed()+xlim(-1,x_len+1)+ylim(-1,y_len+1)+theme_article()+scale_fill_distiller(palette = "Spectral")
```


### Cell-type deconvolution

```{r,eval=FALSE}

pos <- coord
colnames(pos) <- c("x","y")
cd <- round(t(spot_count_layers))
## remove pixels with too few genes
counts <- cleanCounts(cd, min.lib.size = 100)
## feature select for genes
corpus <- restrictCorpus(counts, removeAbove=2.0, removeBelow = 0.01)
## choose optimal number of cell-types
ldas <- fitLDA(t(as.matrix(corpus)), Ks = seq(3, 3, by = 1))
## get best model results
optLDA <- optimalModel(models = ldas, opt = "min")
## extract deconvolved cell-type proportions (theta) and transcriptional profiles (beta)
results <- getBetaTheta(optLDA, perc.filt = 0.05, betaScale = 1000)
deconProp <- results$theta
deconGexp <- results$beta
```

```{r,include=FALSE}

pos <- coord
colnames(pos) <- c("x","y")
cd <- round(t(spot_count_layers))
## remove pixels with too few genes
counts <- cleanCounts(cd, min.lib.size = 100)
## feature select for genes
corpus <- restrictCorpus(counts, removeAbove=2.0, removeBelow = 0.01)
## choose optimal number of cell-types
ldas <- fitLDA(t(as.matrix(corpus)), Ks = seq(3, 3, by = 1))
## get best model results
optLDA <- optimalModel(models = ldas, opt = "min")
## extract deconvolved cell-type proportions (theta) and transcriptional profiles (beta)
results <- getBetaTheta(optLDA, perc.filt = 0.05, betaScale = 1000)
deconProp <- results$theta
deconGexp <- results$beta

evaluation_metric = function(truth,data){
  result_all=list()
  intersect_spot=intersect(rownames(truth),rownames(data))
  truth=truth[intersect_spot,]
  data=data[intersect_spot,]
  
  data=data.frame(data,check.names = F)
  truth=data.frame(truth,check.names=F)
  
  intersect_celltype=intersect(colnames(truth),colnames(data))
  truth=truth[,intersect_celltype]
  data=data[,intersect_celltype]
  
  mse=calc.mse(t(truth), t(data), rsq = FALSE)
  dcor=rep(0,ncol(data))
  diff_all = c()
  for(i in 1:ncol(data)){
    dcor[i]=as.numeric(dcor.test(truth[,i], data[,i], R=200)$statistic)
    diff_all=rbind(diff_all,data.frame(V1=rownames(truth),value=data[,i]-truth[,i],cell_type=colnames(data)[i]))
  }
  result_all[['mse']]=data.table(V1=rownames(truth),value=mse)
  result_all[['rmse']]=data.table(V1=rownames(truth),value=sqrt(mse))
  result_all[['dcor']]=data.table(cell_type=colnames(data),value=dcor)
  result_all[['diff']]=diff_all
  return(result_all)
}

STdeconvolve <- cbind(paste(pos$x,pos$y),deconProp) |> as.data.frame() |> merge(prop_plot,by = "V1")
truth <- STdeconvolve[,c("ct1","ct2","ct3")]
method <- apply(STdeconvolve[,c("1","2","3")],2,as.numeric)
colnames(truth) = c(paste0("celltype",1:3))
colnames(method) = c(paste0("celltype",1:3))
rownames(truth) = STdeconvolve$V1
rownames(method) = STdeconvolve$V1
result_method1=evaluation_metric(as.matrix(truth),as.matrix(method))
```


```{r}
vizAllTopics(deconProp, pos,
             groups = layer_factor, 
             group_cols = rainbow(length(unique(layer_factor))),
             r=2,lwd = 0.5)	

mean(result_method1[['rmse']]$value)
# distance correlation
head(result_method1[['dcor']])
```

### Gene-gene coexpression

#### SpaceX (longleaf R/4.1.0)

```{r,eval=FALSE}
SpaceX_fit <- SpaceX(spot_count_layers,pos,layer_factor,sPMM=FALSE,Post_process = TRUE,numCore = 2)
```

```{r,echo=FALSE}
SpaceX_fit <- readRDS("~/Desktop/ST/SpaceX_4layer.RData")
```

```{r}
corrplot(as.matrix(SpaceX_fit[["Shared_network"]][1:50,1:50]))
```

